---
title: DataSpaceR
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![Build Status](https://travis-ci.org/FredHutch/DataSpaceR.svg?branch=master)](https://travis-ci.org/FredHutch/DataSpaceR)
[![codecov](https://codecov.io/gh/FredHutch/DataSpaceR/branch/master/graph/badge.svg)](https://codecov.io/gh/FredHutch/DataSpaceR/branch/master)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/juyeongkim/DataSpaceR?branch=master&svg=true)](https://ci.appveyor.com/project/juyeongkim/DataSpaceR)


A thin wrapper around [Rlabkey](https://cran.r-project.org/web/packages/Rlabkey/index.html) to access the [CAVD DataSpace](https://dataspace.cavd.org)
database from R. This package simplifies access to the database for R programmers.

It takes advantage of the standardization of the database to hide all 
the [Rlabkey](https://cran.r-project.org/web/packages/Rlabkey/index.html) 
specific code away from the user. Study-specific datasets can be accessed via 
an object-oriented paradigm.


## Installation

The package can be downloaded here and installed like any other R packages or 
installed directly from github using devtools.

```{r, eval=FALSE}
library(devtools)
install_github("FredHutch/DataSpaceR")
```


## netrc

The database is accessed with the user's credentials. A `netrc` file storing 
login and password information is required.

Create netrc file in the computer running R.

* On a UNIX system this file should be named `.netrc` (**dot**`netrc`) 
* On windows it sould be named `_netrc` (**underscore**`netrc`). 
* The file should be located in the users home directory and the permissions
on the file should be unreadable for everybody except the owner. To determine
home directory, run `Sys.getenv("HOME")` in R.

The following three lines must be included in the `.netrc` or `_netrc` file 
either separated by white space (spaces, tabs, or newlines) or commas.

```
machine dataspace.cavd.org
login myuser@domain.com
password supersecretpassword
```

Multiple such blocks can exist in one file.

See [here](https://www.labkey.org/wiki/home/Documentation/page.view?name=netrc) 
for more information about `netrc`.


## Usage

The general idea is that the user creates an instance of an 
`DataSpaceConnection` class. The instance configures itself to connect to 
a specific study, and datasets can be retrieved by name.


### For example:

```{r}
library(DataSpaceR)
study <- connectDS("cvd408")
study
class(study)
```

`connectDS("cvd408")` will create an instance of `cvd408`. The user needs 
credentials stored in a `netrc` file to access the database. 


### Datasets can be listed by:

```{r}
study$availableDatasets
```

which will print names of available datasets.


### Neutralizing antibody dataset (`NAb`) can be retreived by:

```{r}
NAb <- study$getDataset("NAb")
dim(NAb)
colnames(NAb)
```

The package uses a [R6](https://cran.r-project.org/web/packages/R6/index.html) 
class to represent the connection to a study and get around some of R's 
copy-on-change behaviour.
