---
title: "Introduction to DataSpaceR"
author: "Ju Yeong Kim"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to DataSpaceR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = NOT_CRAN,
  eval = NOT_CRAN
)
```


This package provides a thin wrapper around [Rlabkey](https://cran.r-project.org/package=Rlabkey) and connects to the the [CAVD DataSpace](https://dataspace.cavd.org) database, making it easier to fetch datasets from specific studies.


## Configuration

First, go to [DataSpace](https://dataspace.cavd.org) now and set yourself up with an account.

In order to connect to the CAVD DataSpace via `DataSpaceR`, you will need a `netrc` file in your home directory that will contain a `machine` name (hostname of DataSpace), and `login` and `password`. There are two ways to create a `netrc` file.

### Creating a netrc file with `writeNetrc`

On your R console, create a `netrc` file using a function from `DataSpaceR`:

```{r, eval=FALSE}
writeNetrc(
  login = "yourEmail@address.com", 
  password = "yourSecretPassword",
  netrcFile = "/your/home/directory/.netrc" # use getNetrcPath() to get the default path 
)
```

This will create a `netrc` file in your home directory. Make sure you have a valid login and password.

### Manually creating a netrc file

***Alternatively***, you can manually create a netrc file.

* On Windows, this file should be named `_netrc`
* On UNIX/Mac, it should be named `.netrc`
* The file should be located in the user's home directory, and the permissions on the file should be unreadable for everybody except the owner
* To determine your home directory, run `Sys.getenv("HOME")` in R

The following three lines must be included in the `.netrc` or `_netrc` file either separated by white space (spaces, tabs, or newlines) or commas. Multiple such blocks can exist in one file.

```
machine dataspace.cavd.org
login myuser@domain.com
password supersecretpassword
```

See [here](https://www.labkey.org/wiki/home/Documentation/page.view?name=netrc) for more information about `netrc`.


## Initiate a connection

We'll be looking at study `cvd256`. If you want to use a different study, change that string. You can instantiate multiple connections to different studies simultaneously.

```{r connectDS}
library(DataSpaceR)
con <- connectDS()
con
```

The call to `connectDS` instantiates the connection. Printing the object shows where it's connected and the available studies.

```{r availableStudies}
knitr::kable(head(con$availableStudies))
```

`con$availableStudies` shows the available studies in the CAVD DataSpace. Check out [the reference page](https://ropensci.github.io/DataSpaceR/reference/DataSpaceConnection.html) of `DataSpaceConnection` for all available fields and methods.

```{r getStudy}
cvd256 <- con$getStudy("cvd256")
cvd256
```

`con$getStudy` creates a connection to the study `cvd256`. Printing the object shows where it's connected, to what study, and the available datasets.

```{r other-fields}
knitr::kable(cvd256$availableDatasets)
knitr::kable(cvd256$treatmentArm)
```

Available datasets and treatment arm information for the connection can be accessed by `availableDatasets` and `treatmentArm`.


## Fetching datasets

We can grab any of the datasets listed in the connection (`availableDatasets`).

```{r getDataset}
NAb <- cvd256$getDataset("NAb")
dim(NAb)
colnames(NAb)
```

The *cvd256* object is an [`R6`](https://cran.r-project.org/package=R6) class, so it behaves like a true object. Functions (like `getDataset`) are members of the object, thus the `$` semantics to access member functions.

We can get detailed variable information using `getDatasetDescription`.

```{r getVariableInfo}
knitr::kable(cvd256$getDatasetDescription("NAb"))
```

To get only a subset of the data and speed up the download, filters can be passed to `getDataset`. The filters are created using the `makeFilter` function of the `Rlabkey` package.

```{r getDataset-filter}
cvd256Filter <- makeFilter(c("visit_day", "EQUAL", "0"))
NAb_day0 <- cvd256$getDataset("NAb", colFilter = cvd256Filter)
dim(NAb_day0)
```

See `?makeFilter` for more information on the syntax.


## Creating a connection to all studies

To fetch data from multiple studies, create a connection at the project level.

```{r connection-all-studies}
cavd <- con$getStudy("")
```

This will instantiate a connection at the `CAVD` level. Most functions work cross study connections just like they do on single studies.

You can get a list of datasets available across all studies.

```{r connection-all-studies-datasets}
cavd
knitr::kable(cavd$availableDatasets)
```

In all-study connection, `getDataset` will combine the requested datasets. Note that in most cases, the datasets will have too many subjects for quick data transfer, making filtering of the data a necessity. The `colFilter` argument can be used here, as described in the `getDataset` section.

```{r connection-all-studies-dem}
conFilter <- makeFilter(c("species", "EQUAL", "Human"))
human <- cavd$getDataset("Demographics", colFilter = conFilter)
dim(human)
colnames(human)
```

Check out [the reference page](https://ropensci.github.io/DataSpaceR/reference/DataSpaceStudy.html) of `DataSpaceStudy` for all available fields and methods.


## Connect to a saved group

A group is a curated collection of participants from filtering of treatments, products, studies, or species, and it is created in [the DataSpace App](https://dataspace.cavd.org/cds/CAVD/app.view).

Let's say you are using the App to filter and visualize data and want to save them for later or explore in R with `DataSpaceR`. You can save a group by clicking the Save button on the Active Filter Panel.

We can browse available the saved groups or the curated groups by DataSpace Team via `availableGroups`.

```{r availableGroups}
knitr::kable(con$availableGroups)
```

To fetch data from a saved group, create a connection at the project level with a group ID. For example, we can connect to the "NYVAC durability comparison" group which has group ID 220 by `getGroup`.

```{r group-connection}
nyvac <- con$getGroup(220)
nyvac
```

Retrieving a dataset is the same as before.

```{r group-connection-nab}
NAb_nyvac <- nyvac$getDataset("NAb")
dim(NAb_nyvac)
```


## Reference Tables

The followings are the tables of all fields and methods that work on `DataSpaceConnection` and `DataSpaceStudy` objects and could be used as a quick reference.

### `DataSpaceConnection`

| Name | Description |
| --- | --- |
| `availableStudies` | The table of available studies. |
| `availableGroups` | The table of available groups. |
| `getStudy` | Create a `DataSpaceStudy` object by study. |
| `getGroup` | Create a `DataSpaceStudy` object by group. |

### `DataSpaceStudy`

| Name | Description |
| --- | --- |
| `study` | The study name. |
| `group` | The group name. |
| `availableDatasets` | The table of datasets available in the study object. |
| `treatmentArm` | The table of treatment arm information for the connected study. Not available for all study connection. |
| `studyInfo` | Stores the information about the study. |
| `getDataset` | Get a dataset from the connection. |
| `getDatasetDescription` | Get variable information. |


## Session information

```{r session-info}
sessionInfo()
```
