---
title: "Introduction to DataSpaceR"
author: "Ju Yeong Kim"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to DataSpaceR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package provides a thin wrapper around [Rlabkey](https://cran.r-project.org/web/packages/Rlabkey/index.html) and connects to the the [CAVD DataSpace](https://dataspace.cavd.org) database, making it easier to fetch datasets from specific studies.


## Configuration

First, go to [DataSpace](https://dataspace.cavd.org) now and set yourself up with an account.

In order to connect to the CAVD DataSpace via `DataSpaceR`, you will need a `netrc` file in your home directory that will contain a `machine` name (hostname of DataSpace), and `login` and `password`. There are two ways to create a `netrc` file.

### Creating a netrc file with `writeNetrc`

On your R console, create a `netrc` file using a function from `DataSpaceR`:

```{r, eval=FALSE}
DataSpaceR::writeNetrc("yourEmail@address.com", "yourSecretPassword")
```

This will create a `netrc` file in your home directory. Make sure you have a valid login and password.

### Manually creating a netrc file

***Alternatively***, you can manually create a netrc file.

* On Windows, this file sould be named `_netrc`
* On UNIX/Mac, it should be named `.netrc`
* The file should be located in the user's home directory, and the permissions on the file should be unreadable for everybody except the owner
* To determine your home directory, run `Sys.getenv("HOME")` in R

The following three lines must be included in the `.netrc` or `_netrc` file either separated by white space (spaces, tabs, or newlines) or commas. Multiple such blocks can exist in one file.

```
machine dataspace.cavd.org
login myuser@domain.com
password supersecretpassword
```

See [here](https://www.labkey.org/wiki/home/Documentation/page.view?name=netrc) for more information about `netrc`.


## Initiate a connection

We'll be looking at study `cvd256`. If you want to use a different study, change that string. You can instantiate multiple connections to different studies simultaneously.

```{r connectDS}
library(DataSpaceR)
con <- connectDS()
con
```

The call to `connectDS` instantiates the connection. Printing the object shows where it's connected and the available studies.

```{r getStudy}
cvd256 <- con$getStudy("cvd256")
cvd256
```

`con$getStudy` creates a connection to the study `cvd256`. Printing the object shows where it's connected, to what study, and the available datasets.

```{r other-fields}
knitr::kable(cvd256$availableDatasets)
knitr::kable(cvd256$treatmentArm)
```

Available datasets and treatment arm information for the connection can be accessed by `availableDatasets` and `treatmentArm`.


## Fetching datasets

We can grab any of the datasets listed in the connection (`availableDatasets`).

```{r getDataset}
NAb <- cvd256$getDataset("NAb")
dim(NAb)
colnames(NAb)
```

The *cvd256* object is an [`R6`](https://cran.r-project.org/web/packages/R6/) class, so it behaves like a true object. Functions (like `getDataset`) are members of the object, thus the `$` semantics to access member functions.

We can get detailed variable information using `getVariableInfo`.

```{r getVariableInfo}
knitr::kable(cvd256$getVariableInfo("NAb"))
```

To get only a subset of the data and speed up the download, filters can be passed to `getDataset`. The filters are created using the `makeFilter` function of the `Rlabkey` package.

```{r getDataset-filter}
cvd256Filter <- makeFilter(c("visit_day", "EQUAL", "0"))
NAb_day0 <- cvd256$getDataset("NAb", colFilter = cvd256Filter)
dim(NAb_day0)
```

See `?makeFilter` for more information on the syntax.


## Cross-study connection

To fetch data from multiple studies, create a connection at the project level.

```{r cross-connection}
cavd <- con$getStudy("")
```

This will instantiate a connection at the `CAVD` level. Most functions work cross study connections just like they do on single studies.

You can get a list of datasets available across all studies.

```{r cross-connection-print}
cavd
knitr::kable(cavd$availableDatasets)
```

In cross-study connections, `getDataset` will combine the requested datasets. Note that in most cases the datasets will have too many subjects for quick data transfer, making filtering of the data a necessity. The `colFilter` argument can be used here, as described in the `getDataset` section.

```{r cross-connection-dem}
conFilter <- makeFilter(c("species", "EQUAL", "Human"))
human <- cavd$getDataset("Demographics", colFilter = conFilter)
dim(human)
colnames(human)
```


## Connect to a saved group

We can browse available saved groups via `availableGroups`.

```{r availableGroups}
knitr::kable(con$availableGroups)
```

To fetch data from a saved group, create a connection at the project level with a group ID. For example, we can connect to `mice` group which has group ID 216 by `getGroup`.

```{r group-connection}
mice <- con$getGroup(216)
mice
```

Retrieving a dataset is the same as before.

```{r group-connection-nab}
NAb_mice <- mice$getDataset("NAb")
dim(NAb_mice)
```


## Caching

### Connect to `cvd408` study

```{r}
library(pryr)
cvd408 <- con$getStudy("cvd408")

str(cvd408$cache)
object_size(cvd408)
```

### Retrieve NAb dataset

```{r}
ptm <- proc.time()
invisible(cvd408$getDataset("NAb"))
proc.time() - ptm

str(cvd408$cache, max.level = 2)
object_size(cvd408)
```

### Retrieve NAb dataset again

```{r}
ptm <- proc.time()
invisible(cvd408$getDataset("NAb"))
proc.time() - ptm

str(cvd408$cache, max.level = 2)
object_size(cvd408)
```

### Clear cache

```{r}
cvd408$clearCache()

str(cvd408$cache, max.level = 2)
object_size(cvd408)
```


## Session information

```{r session-info}
sessionInfo()
```
