---
title: "A Simple Introduction to the DataSpaceR Package"
author: "Ju Yeong Kim"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package provides a thin wrapper around [Rlabkey](https://cran.r-project.org/web/packages/Rlabkey/index.html) and connects to the the [CAVD DataSpace](https://dataspace-staging.cavd.org) database, making it easier to fetch datasets from specific studies.


## Configuration

In order to connect to DataSpace, you will need a `.netrc` file in your home directory that will contain a `machine` name (hostname of DataSpace), and `login` and `password`. See [here](https://www.labkey.org/wiki/home/Documentation/page.view?name=netrc) for more information.

A netrc file may look like this:

```
machine dataspace-staging.cavd.org
login myuser@domain.com
password supersecretpassword
```

### Set up your netrc file now

Put it in your home directory. If you type:

``` 
ls ~/.netrc
```

at the command prompt, you should see it there. If it's not there, create one now. Make sure you have a valid login and password. If you don't have one, go to [DataSpace](https://dataspace-staging.cavd.org) now and set yourself up with an account.


## Initiate a connection

We'll be looking at study `cvd256`. If you want to use a different study, change that string. The connections have state, so you can instantiate multiple connections to different studies simultaneously.

```{r CreateConnection}
library(DataSpaceR)
cvd256 <- connectDS(study = "cvd256")
cvd256
cvd256$availableDatasets
```

The call to `CreateConnection` instantiates the connection Printing the object shows where it's connected, to what study, and the available datasets.


## Fetching datasets

We can grab any of the datasets listed in the connection (`availableDatasets`).

```{r getDataset}
NAb <- cvd256$getDataset("NAb")
dim(NAb)
colnames(NAb)
```

The *cvd256* object is an [`R6`](https://cran.r-project.org/web/packages/R6/) class, so it behaves like a true object. Functions (like `getDataset`) are members of the object, thus the `$` semantics to access member functions.

We can get detailed variable information using `getVariableInfo`.

```{r getVariableInfo}
str(cvd256$getVariableInfo("NAb"))
```

To get only a subset of the data and speed up the download, filters can be passed to `getDataset`. The filters are created using the `makeFilter` function of the `Rlabkey` package.

```{r getDataset-filter}
library(Rlabkey)
cvd256Filter <- makeFilter(c("visit_day", "EQUAL", "0"))
NAb_day0 <- cvd256$getDataset("NAb", colFilter = cvd256Filter)
dim(NAb_day0)
```

See `?makeFilter` for more information on the syntax.


## Cross-study connection

To fetch data from multiple studies, simply create a connection at the project level.

```{r cross-connection}
con <- connectDS("")
```

This will instantiate a connection at the `Studies` level. Most functions work cross study connections just like they do on single studies.

You can get a list of datasets available accross all studies.

```{r cross-connection-print}
con
con$availableDatasets
```

In cross-study connections, `getDataset` will combine the requested datasets. Note that in most cases the datasets will have too many subjects, making the filtering of the data a necessity. The `colFilter` argument can be used here, as described in the `getDataset` section.

```{r cross-connection-dem}
conFilter <- makeFilter(c("species", "EQUAL", "Human"))
human <- con$getDataset("Demographics", colFilter = conFilter)
dim(human)
colnames(human)
```


## Caching

### Connect to `cvd408` study

```{r}
library(pryr)
cvd408 <- connectDS("cvd408")

str(cvd408$cache)
object_size(cvd408)
```

### Retrieve NAb dataset

```{r}
ptm <- proc.time()
invisible(cvd408$getDataset("NAb"))
proc.time() - ptm

str(cvd408$cache, max.level = 2)
object_size(cvd408)
```

### Retrieve NAb dataset again

```{r}
ptm <- proc.time()
invisible(cvd408$getDataset("NAb"))
proc.time() - ptm

str(cvd408$cache, max.level = 2)
object_size(cvd408)
```

### Clear cache

```{r}
cvd408$clearCache()

str(cvd408$cache, max.level = 2)
object_size(cvd408)
```


## Session information

```{r session-info}
sessionInfo()
```
